Nonparametric Noise models for the Gaussian Process

### 2013-06-27 notes/paper.tex
Update on sampling algorithm 1.

\paragraph{Sampling $z_t$} The conditional distribution of $z_t$ can be written as follows
\begin{eqnarray*}
p(z_t | \bm{K}, \bm{y}, \bm{z}_{\lnot t}, \bm{\theta}) & \propto & \int p(\bm{y} | \bm{f}, \bm{z}_{\lnot t}, z_t, \bm{\theta} ) p(\bm{f} | \bm{K}) d \bm{f}, \\
 & = & \int p(y_t | f_t, \theta_{z_t} ) p(f_t | \bm{y}, \bm{z}_{\lnot t}, \bm{\theta}, \bm{K}) df_t.
\end{eqnarray*}

The key bit is that $p(f_t | \bm{y}, \bm{z}_{\lnot t}, \bm{\theta}, \bm{K})$ is a Gaussian process where every datapoint contributes noise that is dependent on the CRP partition it belongs to.

\begin{equation}
\begin{bmatrix}
\bm{y} \\ f_t
\end{bmatrix}
\sim \mathcal{N}\left(\bm{0},
\begin{bmatrix}
\bm{K}(\bm{x}, \bm{x}) + \bm{\Sigma}& \bm{K}(\bm{x},\bm{x_t}) \\
\bm{K}(\bm{x_t}, \bm{x}) & \bm{K}(\bm{x_t},\bm{x_t})
\end{bmatrix}
\right)
\end{equation}

Where $\bm{\Sigma}$ is a diagonal matrix with on the diagonal $\bm{\Sigma}_{ii} = \theta_{z_{ii}}$.

We know that \cite{gpml} $p(\bm{f} | \bm{K})$ can be represented as a multivariate Gaussian distribution.

%  & = & \int p(y_t | f_t, \theta_{z_t} ) p(f_t | \bm{f}, \bm{y}, \bm{z}_{\lnot t}, z_t, \bm{\theta}) p(\bm{y}_{\lnot t} | \bm{f)_{\lnot t},  \bm{z}_{\lnot t}, \bm{\theta}) p(\bm{f} | \bm{K}) d \bm{f}, \\

\paragraph{Sampling $\theta_n$}

>>>>

\paragraph{Sampling $z_t$} The conditional distribution of $z_t$ can be written as follows
\begin{equation}
p(z_t | \bm{K}, \bm{y}, \bm{z}_{\lnot t}, \bm{\theta}, \alpha) \propto p(z_t | \bm{z}_{\lnot t}, \alpha) p(\bm{y} | \bm{z}, \bm{K}, \bm{\theta}).
\end{equation}

We can compute $p(z_t | \bm{z}_{\lnot t}$ easily using the standard conditional probability for a CRP. Given the noise parameters for every CRP partition are fixed, we can compute the marginal likelihood $p(\bm{y} | \bm{z}, \bm{K}, \bm{\theta})$ analytically. (Equation 2.30 in Carl's book).

NOTE: we need to say something about sampling a new $z_t$, algorithm 8 from Radford Neal's paper should be able to make this happen.

\paragraph{Sampling $\bm{\theta}$}
The condition distribution of $\bm{\theta}$ can be written as follows
\begin{equation}
p(\bm{\theta} | \bm{K}, \bm{y}, \bm{z}, H) \propto p(\bm{\theta} | H) p(\bm{y} | \bm{z}, \bm{K}. \bm{\theta})
\end{equation}
Unfortunately $\bm{\theta}$ influences the marginal likelihood in a complicated way which implies that an analytical expression for the condition distribution on $\bm{\theta}$ is impossible.

\subparagraph{Metropolis-Hastings}
The simplest and probably least efficient method to sample form the conditional distribution of $\bm{\theta}$ is to use Metropolis-Hastings sampling. We create a proposal distribution around $\bm{\theta}$ and can then evaluate the conditional probability up to a proportionality constant by computing the marginal likelihood of the observations.

\subparagraph{Auxiliary Variable}
An alternative method to sample from the conditional distribution on $\bm{\theta}$ is to introduce a latent variable $\bm{f}$ for the Gaussian Process. First, we sample
\begin{equation}
p(\bm{f} | \bm{\theta}, \bm{K}, \bm{y}, \bm{z}, H),
\end{equation}
from a multivariate normal distribution.

Given the latent variable $\bm{f}$, each CRP partition becomes independent, and we can sample the $\bm{\theta}$.
\begin{equation}
p(\bm{\theta} | \bm{f}, \bm{K}, \bm{y}, \bm{z}, H) \propto \prod_k p(\theta_k) \prod_n p(y_n | f_n, \theta_{z_n}).
\end{equation}

NOTE: Can we analytically compute the posterior of $\mu,\sigma$? It's learning a number of Normal-Inverse Gamma distributions where the observations are Gaussians (the GP).
